Temporary Trade Barriers: \\ When Will They End?

### 2016-05-13 presentation.tex
final version for InsTED

%\documentclass{beamer} 
\documentclass[handout]{beamer} 
\usetheme{Ilmenau}
\usepackage{graphicx,verbatim,hyperref}
\usepackage{textpos}

@@

\section{Overview}
\subsection{Preview}
\begin{frame}
\frametitle{The Questions}
\pause
\begin{itemize}[<+->]
	\item How long do deviations from trade agreement tariffs last? 
  \item What are the determinants of these deviations?
		\begin{itemize}
			\item They don't happen for \textit{no} reason at all
			\item Take political economy perspective
		\end{itemize}
	\item 
\end{itemize}
\end{frame}

\begin{frame}{Institutional Detail}
\begin{enumerate}[<+->]
	\item Temporary trade barriers have potential to be renewed
		\begin{itemize}
			\item For Anti-dumping (AD) duties, initial five year term, renewal for five years
			\item For Safeguards, four then four; much rarer than AD
		\end{itemize}
	\item Authorized by WTO
	\item Very little in the agreements to guide the process
	\item WTO litigation has given governments a LOT of latitude on renewals
	\item Conclusion
\end{enumerate}
\end{frame}

\begin{frame}{Preview of Results}
\begin{itemize}[<+->]
	\item H
		\begin{itemize}
			\item R
		\end{itemize}
	\item E
		\begin{itemize}
			\item P
			\item R
			\item T
		\end{itemize}
	\item T
\end{itemize}
\end{frame}

@@

\subsection{The Players}
\begin{frame}
\frametitle{``Government''}
Renewal decision determined by complex process including DOC, ITC, pressure via other political bodies. Reduced form:
\[
  W_G = \mathit{CS}_X(\tau) + \ga(e,\ta) \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]

\begin{itemize}
	\item $\mathit{CS_i(\cdot)}$: consumer surplus
	\item $\pi_X(\tau)$: profits of import-competing industry
	\item $\pi_Y(\tau^*)$: profits of exporting industry
	\item $\mathit{TR}(\tau)$: tariff revenue
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Lobby}
\vskip.2in
Lobby chooses effort to maximize:
\[
  \left\{1-\Pr\left[ \text{AD Renewal}\right]\right\} \ \pi(\tau^a) + \Pr\left[ \text{AD Renewal} \right]  \pi(\tau^{\mathit{ad}})  - e
\]

\vskip.1in
\begin{itemize}
	\item $e$: Lobbying effort
	\item $\tau^a$: home import tariff under trade agreement
	\item $\tau^{\textit{ad}}$: home import tariff equivalent under anti-dumping duties
\end{itemize}

@@

\begin{frame}{What's this uncertainty about?}
\pause
Lobby must be able to trigger the original AD duty
\begin{itemize}
	\item \textit{Can} think of non-adherence to MFN as eqm path dispute
	\item It could be justified, not motivated by politics/rent-seeking
	\item But still, firms lobby for duties they don't get
\end{itemize}

@@

[<+->]
\begin{frame}{Exogenous Shifts in $\ga(e,\ta)$}

>>>>

\documentclass{beamer} 
%\documentclass[handout]{beamer} 
\usetheme{Ilmenau}
\usepackage{graphicx,verbatim,hyperref}
\usepackage{textpos}

@@

\section{Overview}
\subsection{Preview}
\begin{frame}
\frametitle{The Questions}
\pause
\begin{enumerate}[<+->]
	\item How long do deviations from trade agreement tariffs last? 
  \item What are the determinants of renewals?
		\begin{itemize}
			\item They don't happen for \textit{no} reason at all
		\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}{Institutional Detail}
\begin{itemize}[<+->]
	\item Temporary trade barriers have potential to be renewed
		\begin{itemize}
			\item For Anti-dumping (AD) duties, initial five year term, renewal for five years
			\item For Safeguards, four then four; much rarer than AD
		\end{itemize}
	\item Authorized by WTO
	\item Very little in the agreements to guide the process
		\begin{itemize}
			\item WTO litigation gives gov'ts a LOT of latitude
		\end{itemize}
	\item In U.S., for AD:
		\begin{itemize}
			\item DOC initiates, determines whether dumping would continue
			\item ITC determines whether injury would recur / continue
		\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Preview of Results}

\pause
The probability that an AD duty gets renewed
\pause
\begin{itemize}[<+->]
	\item Increases in lobbying effort
	\item Decreases in the MFN tariff
	\item Is invariant to the trading partner's tariff
	\item Increases in the profitability of the import-competing sector
	\item Increases in the strength of the lobby
	\item May be concave in the AD duty
\end{itemize}
\end{frame}

@@

\subsection{The Players}
\begin{frame}
\frametitle{``Government''}
\pause
Renewal decision determined by complex process including DOC, ITC, pressure via other political bodies. Reduced form:
\pause
\[
  W_G = \mathit{CS}_X(\tau) + \ga(e,\ta) \pi_X(\tau) + \mathit{CS}_Y(\tau^*) + \pi_Y(\tau^*) + \mathit{TR}(\tau)
\]

\pause
\begin{itemize}[<+->]
	\item $\mathit{CS_i(\cdot)}$: consumer surplus
	\item $\pi_X(\tau)$: profits of import-competing industry
	\item $\pi_Y(\tau^*)$: profits of exporting industry
	\item $\mathit{TR}(\tau)$: tariff revenue
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Lobby}
\vskip.2in
\pause
Lobby chooses effort to maximize:
\[
  \left\{1-\Pr\left[ \text{AD Renewal}\right]\right\} \ \pi(\tau^a) + \Pr\left[ \text{AD Renewal} \right]  \pi(\tau^{\mathit{ad}})  - e
\]

\vskip.1in
\pause
\begin{itemize}[<+->]
	\item $e$: Lobbying effort
	\item $\tau^a$: home import tariff under trade agreement
	\item $\tau^{\textit{ad}}$: home import tariff equivalent under anti-dumping duties
\end{itemize}

@@

\begin{frame}{What's this uncertainty about?}
\pause
Lobby must be able to trigger the original AD duty
\begin{itemize}[<+->]
	\item \textit{Can} think of non-adherence to MFN as eqm path dispute
	\item It could be justified, not motivated by politics/rent-seeking
	\item But still, firms lobby for duties they don't get
\end{itemize}

@@

\begin{frame}{Exogenous Shifts in $\ga(e,\ta)$}
